---
title: "populating attendance worksheets"
author: "John Little"
date:  "`r sys.Date()`"
---

This is a beta script (Jan. 19. 2021).  Not in production.

This script will scrape LibCal and generate a comprehensive data frame of all the workshops listed at LibCal.  (Do this at the beginning of the semester b/c LibCal API will not present past events.)  Then there is data wrangling.  Then using `googleshets4` and the `purrr::walk` function to rowwise iterate the data frame and make an attendance sheet for each workshop found at LibCal.  

Note: The `googlesheets4` library requires realtime interaction with the RStudio console to authenticate the connection Google Drive.

[spring master list - 2021 - manual](https://docs.google.com/spreadsheets/d/1tJ_4D-sEm8kPt8YIZ0_m_9RpQ2-XjOIbs6mHis9oRwE/edit#gid=0)  
[example attendance blank sheet](https://docs.google.com/spreadsheets/d/1zbuWbnOexPq9KkRpv7oRnHqF-pCMSS7aegm3JjhwXkU/edit#gid=1561693836)

```{r}
# This code uses rvest to pull in the current list of DVS workshops
Sys.setenv(TZ="America/New_York")
library(rvest)
library(tidyverse)
library(lubridate)
library(googlesheets4)
```

## Functions

```{r}
get_workshop_id <- . %>% 
  mutate(workshop_id = str_extract(registration, "(?<=event/)\\d+")) 

```


## Ingest data

### column headings 

make a vector of column headings that will become the headings of the google sheests attendance records

```{r}
blank_df <- tibble("Attended (x or blank)" = NA,
                "Registration Status (Registered, Waitlist, or Walk-in)"  = NA,
                # "Workshop Date",
                # "Workshop Name",
                # "WorkshopID",
                "First Name" = NA,
                "Last Name" = NA,
                "Email" = NA,
                "Booking Made" = NA,
                "Attendance" = NA,
                "Are you affiliated with Duke University, Duke Medical Center, DCRI, or another Duke group?" = NA,
                "Academic Status (or other)" = NA,
                "Discipline or Affiliation" = NA,
                "Institutes, Initiatives, or Program Affiliation" = NA,
                "Where did you hear about this event?" = NA,
                "Have you consulted with Data and Visualization Services before this workshop?" = NA,
                "Would you like to receive more information about DVS events and training?" = NA,
                "I'm interested in registering for this workshop because:" = NA,
                "If other, describe below:" = NA)
```

###  scrape webpage

```{r}
dvs_cal <- read_html("https://api3.libcal.com/api_events.php?iid=971&m=upc&cid=3819&c=&d=25858&l=50&target=_blank")
# dvs_cal
```

Using rvest functions to select data via CSS

```{r}
ntitle <- html_nodes(dvs_cal, ".s-lc-ea-ttit a")
ndate  <- html_nodes(dvs_cal, ".s-lc-ea-tdat td:nth-child(2)")
npresenter <- html_nodes(dvs_cal, ".s-lc-ea-tpre td:nth-child(2)")
ntime <- html_nodes(dvs_cal, ".s-lc-ea-ttim td:nth-child(2)")
nlocation <- html_nodes(dvs_cal, ".s-lc-ea-tloc td:nth-child(2)")
#  The description nodes vary a great deal (everyone describes their
#  workshop in different ways).  The code here is likely only going
#  to pull the first paragraph.  Also, note that it has trouble 
#  with people who put in heavily styled code chunks in their 
#  descriptions
ndescription <- html_nodes(dvs_cal, "p:nth-child(1)")
nregistration <- html_nodes(dvs_cal, ".s-lc-ea-treg a")
```

Using rvest functions parse the HTML into vectors

```{r}
title <- html_text(ntitle)
date <- html_text(ndate)
presenter <- html_text(npresenter)
time <- html_text(ntime)
location <- html_text(nlocation)
description <- html_text(ndescription)
registration <- html_attr(nregistration, "href")
```


Clean the description field which gets funky if someone uses too much formatting in the SpringShare::LibCal web-app.  (I'm looking at you, John L.)

```{r}
description <- tibble(description) %>% 
    filter(str_detect(description, "install.packages", negate = TRUE))  %>% 
    filter(str_detect(description, "Introductory familiarity with R \\(consider attending one of our", 
                      negate = TRUE))  %>% 
    filter(description != "A GitHub account") %>%
    pull(description)
# description
```

Make a summary tibble out of the data

```{r}
results_df <- tibble(date, presenter, registration, time, title, description)
results_df
```

Wrangle the data and data types so that the dates are actionable.  Use lubridate to handle the date/time issues.  

```{r}
dvs_cal_tbl <- results_df %>%
  separate(time, c("start_time", "end_time"), sep = "-") %>%
  mutate(date_time_start = force_tz(parse_date_time(str_c(date, start_time, sep = ","), "OmdYIMp"), "America/New_York")) %>%
  mutate(date_time_end  = force_tz(parse_date_time(str_c(date, end_time,   sep = ","), "OmdYIMp"), "America/New_York")) %>%
  get_workshop_id  %>%
  mutate(duration = date_time_end - date_time_start) %>%
  mutate(workshop_duration_minutes = as.numeric(duration) * 60) %>% 
  mutate(date = as_date(date_time_start)) %>% 
  select(
    workshop_id,
    date,
    title,
    presenter,
    workshop_duration_minutes,
    date_time_start,
    date_time_end,
    description,
    registration,
    location
  ) %>%
  mutate(location = NA) %>% 
  mutate(theme = NA) %>%
  mutate(theme2 = NA) %>%
  mutate(cname = NA) %>%
  mutate(status = NA) 

dvs_cal_tbl
```    
  
## Make an example attendance sheet

```{r}
attendance_template_tbl <- dvs_cal_tbl %>% 
  bind_cols(blank_df) %>% 
  rename(`Workshop Date` = date) %>%
  select(15:16, `Workshop Date`,
         `Workshop Name` = title,
         `WorkshopID` = workshop_id,
         17:30
         )
attendance_template_tbl 
```

## Write google sheets

### Master Workshop List

**First**, write the _Master Workshop List_ to google drive.

This automates the list of workshops identified at the beginning of a semester.  It is derived from a webscrape of springshare::libcal at the beginning of the semester.  If this is run later in the semester, it only include the workshops from the date of the web-scrape and moving forward.  i.e. It will not include January and February workshops if run in March.

```{r}
workshop_summary_list <- dvs_cal_tbl %>% 
  # get_workshop_id() %>% 
  # relocate(workshop_id) %>% # move `workshop_id` to first column position
  # relocate(title, .after = date)
  rename(worshop_begins = date_time_start) %>% 
  rename(workshop_ends = date_time_end) 
```


The following code chunks are manually disabled because they write files to google drive.  This is "belt and suspenders" caution.  To enable the code-blocks, remove the carriage-return at the end of the third backtick and before the first open curly brace.

```
{r}
current_date <- Sys.Date()
pid <- Sys.getpid()
gs_title <- str_c(current_date,pid, "Master Workshop List", sep = "-")

gs4_create(gs_title, sheets = list(semester_workshops = workshop_summary_list))
```

### iterate over all the workshops and write to Google Drive

**Second**, make one blank attendance template for each workshop. Write each blank template to Google Drive.

```
{r}
workshop_list <- attendance_template_tbl %>% 
   #   slice(n = 23:27) %>% 
   rowwise() %>%
   group_map(tibble)

walk(
  workshop_list,
  ~ gs4_create(.$`Workshop Name`, sheets = list(attendance = .))
)
```

### Sleep / Pause

use `Sys.sleep()` or 
```
{r}
walk(
  seq_len(nrow(band_members)),
  ~ {
    f <- slice(band_members, .)
    gs4_create(f$name,
      sheets = list(sname = f))
    Sys.sleep(5)  # pause for 5 seconds between "write"
  }
)
```



